---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Relational data with dplyr

```{r}
library(dplyr, warn.conflicts = FALSE)
```

[Relational data](https://r4ds.had.co.nz/relational-data.html):

* A collection of related tables of data.

```{r}
companies <- tribble(
  ~companies_id,                                 ~information,
              1,    "alpha sells solar panels and wind mills",
              2, "beta sells steel and installs solar panels",
)

categories <- tribble(
  ~companies_id,       ~sector,
              1,      "energy",
              2,  "metallurgy",
              2,      "energy",
              3, "agriculture",
)
```

* The relations are always defined,  between a pair of tables, by the keys.

```{r}
key  <- "companies_id"

companies |>
  left_join(categories, by = key)
```

```{r echo = FALSE, message=FALSE}
library(dm)

database <- dm(companies, categories) |>
  dm_add_pk(companies, companies_id) |>
  dm_add_fk(categories, companies_id, companies)

database |> 
  dm_draw()
```

### Mutating joins

Add new variables to one data frame from matching observations in another.

--

[Inner join](https://r4ds.had.co.nz/relational-data.html#mutating-joins): Drop
unmatched rows.

```{r}
companies |>
  inner_join(categories, by = "companies_id")
```

[Outter joins](https://r4ds.had.co.nz/relational-data.html#outer-join):

* Keep all rows in `x` (most common).

```{r}
companies |>
  left_join(categories, by = "companies_id")
```

* Keep all rows in `y`

```{r}
companies |>
  right_join(categories, by = "companies_id")
```

* Keep all rows in `x` and `y`

```{r}
companies |>
  full_join(categories, by = "companies_id")
```

* [Duplicate keys](https://r4ds.had.co.nz/relational-data.html#join-matches)

```{r}
# Duplicate keys in one table
companies2 <- companies |>
  bind_rows(tibble(companies_id = 1L, information = "abc"))

companies2 |>
  left_join(categories)

# Duplicate keys in both tables
categories2 <- categories |>
  bind_rows(tibble(companies_id = 1L, sector = "xyz"))

companies2 |>
  left_join(categories2)
```

### Filtering joins

Affect observations, not variables:

* Keep all observations in `x` that have a match in `y`.

```{r}
categories |> 
  semi_join(companies)
```

* Drop all observations in `x` that have a match in `y`.

```{r}
categories |> 
  anti_join(companies)
```

### Validating keys to avoid [join problems](https://r4ds.had.co.nz/relational-data.html#join-problems)

* All values of a primary key should be unique.

```{r}
companies |>
  count(companies_id) |>
  filter(n > 1)

# (!) The `companies_id` column is NOT a primary key of `categories`
categories |>
  count(companies_id) |>
  filter(n > 1)
```

* No value of a primary key should be missing.

```{r}
companies |>
  filter(is.na(companies_id))
```

* All values of a foreign key should match a value of a primary key.

```{r}
# ! Problem: Invalid foreign key
anti_join(categories, companies, by = "companies_id")
```
