---
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Using a Relational Database Management System (RDBMS)

Packages.

```{r}
library(dm, warn.conflicts = FALSE)
library(dplyr, warn.conflicts = FALSE)
library(DBI)
library(RSQLite)

library(here)
```

The dm package make it easy to deploy a data model to a Relational Database Management System (RDBMS).

### Deploy a data model to a RDBMS

If you have an RDBMS (an SQLite file) ...

```{r}
# See arguments to dbConnect() or the relevant driver at ?RSQLite::RSQLite()
database_file <- here("04_rdbms", "database.sqlite")
connection <- DBI::dbConnect(RSQLite::SQLite(), dbname = database_file)
connection
```

... and you have a data model ...

```{r}
# styler: off
companies <- tribble(

  ~companies_id,                                 ~information,
              1,    "alpha sells solar panels and wind mills",
              2, "beta sells steel and installs solar panels",
)

categories <- tribble(
  ~companies_id,       ~sector,
              1,      "energy",
              2,  "metallurgy",
              2,      "energy",
)
# styler: on

dm <- dm(companies, categories)
```

... then you can easily deploy the data model to the RDBMS with `dm::copy_dm_to()`:

* Defaults to creating temporary tables
* Sets primary key constraints on all databases
* Sets foreign key constraints are set on MSSQL and Postgres databases.

```{r}
copy_dm_to(connection, dm, temporary = FALSE)
```

Once you're finish working with the database, you should disconnect.

```{r}
DBI::dbDisconnect(connection)
```

### Using the RDBMS

Connect to the RDBMS.

```{r}
database_file <- here("04_rdbms", "database.sqlite")
connection <- RSQLite::SQLite() |> DBI::dbConnect(dbname = database_file)
```

Use it with [dplyr](https://dplyr.tidyverse.org/) (and [dbplyr](https://dbplyr.tidyverse.org/) in the backend) (meetup 1).

```{r}
connection |> DBI::dbListTables()

companies <- connection |> dplyr::tbl("companies")
categories <- connection |> dplyr::tbl("categories")

companies |> left_join(categories, by = "companies_id")
```

Use it with [dm](https://github.com/cynkra/dm) (meetup 2).

```{r}
# Helper
companies_dm <- function(connection) {
  dm(
    companies = connection |> tbl("companies"),
    categories = connection |> tbl("categories")
  ) |>
    dm_add_pk(companies, companies_id) |>
    dm_add_fk(categories, companies_id, companies)
}

dm <- companies_dm(connection)
dm

dm |> dm_flatten_to_tbl(.start = categories)
```

```{r}
connection |> DBI::dbDisconnect()
```

```{r include=FALSE}
# Cleanup
fs::file_delete(database_file)
```
